Visualizing probability distributions across bivariate cyclic temporal granularities

### 2020-03-12 main.Rmd
include interchangeability mapping

### Interaction {#sec:vis-interaction}

We will discuss the effect of the interaction of cyclic granularities in this section.<!--In Section \ref{sec:synergy}, we discussed how pairs of granularities can have empty combinations either due to structure of calendar, event location or duration or due to the way they are built. In this section, we will see how these empty combinations affect the visualization when a dependent variable is plotted against these granularities.--> For illustration, distribution of half-hourly electricity consumption of Victoria is plotted across different time granularities in each of the panel in
\autoref{fig:allFig}. \autoref{fig:allFig} (a) shows the letter value plot across days of the month faceted by months like January, March and December. <!---\autoref{fig:allFig} (c) shows box plot  across days of the year by the 1st, 15th, 29th and 31st days of the month. \autoref{fig:allFig} (d) showing violin plot across days of the month faceted by week of the month. \autoref{fig:allFig} (e), variations across week of the year conditional on week of the month can be observed through a ridge plot and \autoref{fig:allFig}---> (b) shows decile plots across day of the year and month of the year.

Clearly, in \autoref{fig:allFig}, we observe that choice (a) of cyclic time granularities work but (b) does not work. In (b) we can not compare the distributions across different facets becuase every day-of-year cannot occur for every month. For example, 21st day of the year can never correspond to the month March or December of any year.
<!---In \autoref{fig:allFig} (c), there will be no observations for some combinations day-of-month and day-of-year. In particular, the 1st day of the month can never correspond to 2nd, 3rd or 4th day of the year. On the contrary, for \autoref{fig:allFig}---> In (a) we will not have any combinations because every day-of-week can occur in every month-of-year. <!---Thus the graphs that don't work are those where many of the combination sets are empty.--->

In other words, if there are levels of cyclic granularities plotted across x-axis which are not spanned by levels of cyclic granularities plotted across facets or vice versa, we will have empty sets leading to potential ineffective graphs. We hypothesize that the synergy of these cyclic granularities are thus playing a role while deciding if the resulting plot would be a good candidate for exploratory analysis.

We redefine harmony and clashes as follows: harmonies are pairs of granularities that do not contain emptry combinations and thus aid exploratory data analysis, and clashes are pairs that do not aid the process of exploratory analysis because they contain empty combinations. As a result, we should avoid plotting clashes.<!---as they hinder the exploratory process by having missing combinations of time granularities in each panel.--->

@@

# change all of these to gravitas function
library(gravitas)
library(tsibbledata)
library(dplyr)
library(ggplot2)

@@

VIC %>%
  create_gran("month_year") %>%
  filter(month_year %in% c("Jan", "Mar", "July", "Dec")) %>%
  prob_plot("month_year", "day_week",
    response = "Demand",
    plot_type = "lv"
  ) +
  ylab("Electricity Demand [KWh]") +
  ggtitle("(a) Letter value plot")

@@

VIC %>%
  create_gran("month_year") %>%
  filter(month_year %in% c("Jan", "Jul", "Nov")) %>%
  prob_plot("month_year",
    "day_year",
    response = "Demand",
    plot_type = "quantile",
    quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9),
    symmetric = FALSE
  ) +
  # scale_x_discrete(breaks = seq(1, 336, 60)) +
  ylab("") +
  xlab("Day of the Year") + theme(
    legend.position = "bottom",
    strip.text = element_text(
      size = 7,
      margin = margin()
    )
  ) +
  ggtitle("(b) Decile plot")

@@

We will discuss the effect of the mapping of cyclic granularities in this section. When we consider data sets of the form <$C_1$, $C_2$, $v$> with $C_1$ mapped to $x$ position and $C_2$ to facets $A_i$'s are placed in close proximity and each $B_j$ represent a group/facet. Gestalt theory suggests that when items are placed in close proximity, people assume that they are in the same group because they are close to one another and apart from other groups. Hence, in this case $A_i$'s are compared against each other within each group.<!-- When $C_2$ is mapped to facets and $C_1$ is mapped to $x$ axis, comparisons would ideally happen across all $B_j$ within each $A_i$-->. With the mapping of $C1$ and $C_2$ reversed, emphasis will shift to different behavior of the variables.

>>>>

### Synergy of cyclic granularities {#sec:vis-interaction}

If there are levels of cyclic granularities plotted across x-axis which are not spanned by levels of cyclic granularities plotted across facets or vice versa, we will have empty sets leading to potential ineffective graphs. We hypothesize that the synergy of these cyclic granularities are thus playing a role while deciding if the resulting plot would be a good candidate for exploratory analysis. Harmonies are pairs of granularities that do not contain emptry combinations and thus aid exploratory data analysis, and clashes are pairs that do not aid the process of exploratory analysis because they contain empty combinations. As a result, we should avoid plotting clashes.

<!--In Section \ref{sec:synergy}, we discussed how pairs of granularities can have empty combinations either due to structure of calendar, event location or duration or due to the way they are built. In this section, we will see how these empty combinations affect the visualization when a dependent variable is plotted against these granularities.--> For illustration, distribution of half-hourly electricity consumption of Victoria is plotted across different time granularities in each of the panel in \autoref{fig:allFig}. \autoref{fig:allFig} (a) shows the letter value plot across days of the month faceted by months like January, March and December. This plot works because every day-of-week can occur in every month-of-year.

<!---\autoref{fig:allFig} (c) shows box plot  across days of the year by the 1st, 15th, 29th and 31st days of the month. \autoref{fig:allFig} (d) showing violin plot across days of the month faceted by week of the month. \autoref{fig:allFig} (e), variations across week of the year conditional on week of the month can be observed through a ridge plot and \autoref{fig:allFig}(b) shows decile plots across day of the year and month of the year.--->

<!-- Clearly, in \autoref{fig:allFig}, we observe that  -->

<!--In (b) we can not compare the distributions across different facets becuase every day-of-year cannot occur for every month. For example, 21st day of the year can never correspond to the month March or December of any year.-->
<!---In \autoref{fig:allFig} (c), there will be no observations for some combinations day-of-month and day-of-year. In particular, the 1st day of the month can never correspond to 2nd, 3rd or 4th day of the year. On the contrary, for \autoref{fig:allFig}--->  <!---Thus the graphs that don't work are those where many of the combination sets are empty.--->

<!---as they hinder the exploratory process by having missing combinations of time granularities in each panel.--->

@@

library(gravitas)
library(tsibbledata)
library(dplyr)
library(ggplot2)

@@

VIC %>%
  create_gran("month_year") %>%
  filter(month_year %in% c("Jan", "Mar", "Dec")) %>% 
  create_gran("day_week") %>%
  filter(day_week %in% c("Sat", "Sun", "Mon")) %>%
  prob_plot("month_year", "day_week",
    response = "Demand",
    plot_type = "lv"
  ) +
  ylab("Electricity Demand [KWh]") +
  ggtitle("(a) Letter value plot with month-of-year facet")

@@

# VIC %>%
#   create_gran("month_year") %>%
#   filter(month_year %in% c("Jan", "Jul", "Nov")) %>%
#   prob_plot("month_year",
#     "day_year",
#     response = "Demand",
#     plot_type = "quantile",
#     quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9),
#     symmetric = FALSE
#   ) +
#   # scale_x_discrete(breaks = seq(1, 336, 60)) +
#   ylab("") +
#   xlab("Day of the Year") + theme(
#     legend.position = "bottom",
#     strip.text = element_text(
#       size = 7,
#       margin = margin()
#     )
#   ) +
#   ggtitle("(b) Decile plot")

VIC %>%
  create_gran("month_year") %>%
  filter(month_year %in% c("Jan", "Mar", "Dec")) %>% 
  create_gran("day_week") %>%
  filter(day_week %in% c("Sat", "Sun", "Mon")) %>%
  prob_plot("day_week",
    "month_year",
    response = "Demand",
    plot_type = "lv"
  ) +
  ylab("") +
  xlab("Day of the Year") +
  ggtitle("(b) Letter value plot with day-of-week facet")

@@

We will discuss the effect of the mapping of cyclic granularities in this section. When we consider data sets of the form <$C_1$, $C_2$, $v$> with $C_1$ mapped to $x$ position and $C_2$ to facets $A_i$'s are placed in close proximity and each $B_j$ represent a group/facet. Gestalt theory suggests that when items are placed in close proximity, people assume that they are in the same group because they are close to one another and apart from other groups. Hence, in this case $A_i$'s are compared against each other within each group.<!-- When $C_2$ is mapped to facets and $C_1$ is mapped to $x$ axis, comparisons would ideally happen across all $B_j$ within each $A_i$-->. With the mapping of $C1$ and $C_2$ reversed, emphasis will shift to different behavior of the variables. \autoref{fig:allFig} (a) shows the letter value plot across days of the week faceted by months like January, March and December and \autoref{fig:allFig} (b) shows letter value of across months faceted by days. The first one helps us to compare different days of the weeks within a month and the second one helps to compare different months for each day of the week.
